---
title: "Exercises for class 1"
output: html_document
date: "13-02-24"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Exercises for Class 1

## Supporting code

Below is code from Chapter 2 of the book that you can use to do the exercises. Not all the code you require is here though. Make sure to not rush it and take time to play around with each function. :))

```{r}
# Grid Approximation

# define grid
p_grid = seq(from = 0, to = 1, length.out=20)
p_grid

# define prior
prior = rep(1, 20)

# compute likelihood at each value in grid
likelihood = dbinom(6, size = 9, prob = p_grid)

# compute product of likelihood and prior
unstd.posterior = likelihood * prior

# standardize the posterior so it sums to 1
posterior = unstd.posterior / sum(unstd.posterior)

# plot the distribution
plot(p_grid, posterior, type = "b",
     xlab = "probability of water", ylab = "posterior_probability")
mtext("20 points")

```

## Exercises

### Easy

2E1. 

(2) Pr(rain|Monday)

2E2.

(3) The probability that it is Monday, given that it is raining.

2E3.

(1) Pr(Monday|rain)

2E4.

The probability of water being 0.7 means that if you were to toss the globe e.g. 100 times, your finger would land on water roughly 70 times; but even if you throw it a different amt of times, you would still land on water roughly 70& of the time. One could also say a probability is a distribution of outcomes, with the mean being the "probability".

### Medium

2M1.
Recall the globe tossing model from the chapter. Compute and plot the grid approximate posterior distribution for each of the following sets of observations. In each case, assume a uniform prior for p. 
(1) W, W, W 
(2) W, W, W, L 
(3) L, W, W, L, W, W, W

```{r}
#define the grid
p_grid = seq(from = 0, to = 1, length.out=20)

#define the prior
prior <- rep(1,20) # if i multiply prior with likelihood, posterior stays the same; equal chance of any amount of success, i.e. equal chance of earth to be covered in any amount of water

#compute likelihood (probability!) at each value in grid
likelihood <- dbinom(3, size=3, prob=p_grid) # if i have 3 trials and 3 waters, whats the probability that earth is p_grid amount covered with water, p(b|a)? what is the probability of getting 3 out of 3 successful trials in total, given our probability of success is 0.7? how surprising is it (inversely)? prob=p_grid means probability for every value of p grid 

#compute product of likelihood and prior
unstd.posterior <- likelihood*prior

#standardize posterior, so it sums to 1
posterior <- unstd.posterior / sum(unstd.posterior)

# plot the distribution
plot(p_grid, posterior, type = "b",
     xlab = "coverage of water", ylab = "posterior_probability")
mtext("20 points")


prior
likelihood
```

2M2.

```{r}
# define grid
p_grid = seq(from = 0, to = 1, length.out=20)
p_grid

# define prior
prior = c(rep(0, 10),rep(1,10))

# compute likelihood at each value in grid
likelihood = dbinom(5, size = 7, prob = p_grid)

# compute product of likelihood and prior
unstd.posterior = likelihood * prior # because you multiply 0 0 0 0 1 1 1 1 1 with the actual values, the 0 will be 0 but the 1 wont change!!!

# standardize the posterior so it sums to 1
posterior = unstd.posterior / sum(unstd.posterior)

# plot the distribution
plot(p_grid, posterior, type = "b",
     xlab = "probability of water", ylab = "posterior_probability")
mtext("20 points")
```

2M3.
Suppose there are two globes, one for Earth and one for Mars. The Earth globe is 70% covered in water. The Mars globe is 100% land. Further suppose that one of these globes—you don’t know which—was tossed in the air and produced a “land” observation. Assume that each globe was equally likely to be tossed. Show that the posterior probability that the globe was the Earth, conditional on seeing “land” (Pr(Earth|land)), is 0.23.
```{r}

# Use the Bayes formula!

```

For the exercises below, I highly suggest you to grab a piece of paper and try to solve the problems in a 'visual' way.

2M4.

```{r}

# Write your probability here.

```

2M5.

```{r}

# Write your probability here.

```

2M6.

```{r}

# Write your probability here.

```
