---
title: "Exercises for class 1"
output: html_document
date: "13-02-24"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Exercises for Class 1

## Supporting code

Below is code from Chapter 2 of the book that you can use to do the exercises. Not all the code you require is here though. Make sure to not rush it and take time to play around with each function. :))

```{r}
# Grid Approximation

# define grid
p_grid = seq(from = 0, to = 1, length.out=20)
p_grid

# define prior
prior = rep(1, 20)

# compute likelihood at each value in grid
likelihood = dbinom(6, size = 9, prob = p_grid)

# compute product of likelihood and prior
unstd.posterior = likelihood * prior

# standardize the posterior so it sums to 1
posterior = unstd.posterior / sum(unstd.posterior)

# plot the distribution
plot(p_grid, posterior, type = "b",
     xlab = "probability of water", ylab = "posterior_probability")
mtext("20 points")

```

## Exercises

### Easy

2E1. 

(2) Pr(rain|Monday)

2E2.

(3) The probability that it is Monday, given that it is raining.

2E3.

(1) Pr(Monday|rain)

2E4.

The probability of water being 0.7 means that if you were to toss the globe e.g. 100 times, your finger would land on water roughly 70 times; but even if you throw it a different amt of times, you would still land on water roughly 70& of the time. One could also say a probability is a distribution of outcomes, with the mean being the "probability".

### Medium

2M1.
Recall the globe tossing model from the chapter. Compute and plot the grid approximate posterior distribution for each of the following sets of observations. In each case, assume a uniform prior for p. 
(1) W, W, W 
(2) W, W, W, L 
(3) L, W, W, L, W, W, W

```{r}
#define the grid
p_grid = seq(from = 0, to = 1, length.out=20)

#define the prior
prior <- rep(1,20) # if i multiply prior with likelihood, posterior stays the same; equal chance of any amount of success, i.e. equal chance of earth to be covered in any amount of water

#compute likelihood (probability!) at each value in grid
likelihood <- dbinom(3, size=3, prob=p_grid) # if i have 3 trials and 3 waters, whats the probability that earth is p_grid amount covered with water, p(b|a)? what is the probability of getting 3 out of 3 successful trials in total, given our probability of success is 0.7? how surprising is it (inversely)? prob=p_grid means probability for every value of p grid 

#compute product of likelihood and prior
unstd.posterior <- likelihood*prior

#standardize posterior, so it sums to 1
posterior <- unstd.posterior / sum(unstd.posterior)

# plot the distribution
plot(p_grid, posterior, type = "b",
     xlab = "coverage of water", ylab = "posterior_probability")
mtext("20 points")


prior
likelihood
```

2M2.

```{r}
# define grid
p_grid = seq(from = 0, to = 1, length.out=20)
p_grid

# define prior
prior = c(rep(0, 10),rep(1,10))

# compute likelihood at each value in grid
likelihood = dbinom(5, size = 7, prob = p_grid)

# compute product of likelihood and prior
unstd.posterior = likelihood * prior # because you multiply 0 0 0 0 1 1 1 1 1 with the actual values, the 0 will be 0 but the 1 wont change!!!

# standardize the posterior so it sums to 1
posterior = unstd.posterior / sum(unstd.posterior)

# plot the distribution
plot(p_grid, posterior, type = "b",
     xlab = "probability of water", ylab = "posterior_probability")
mtext("20 points")
```

2M3.
Suppose there are two globes, one for Earth and one for Mars. The Earth globe is 70% covered in water. The Mars globe is 100% land. Further suppose that one of these globes—you don’t know which—was tossed in the air and produced a “land” observation. Assume that each globe was equally likely to be tossed. Show that the posterior probability that the globe was the Earth, conditional on seeing “land” (Pr(Earth|land)), is 0.23.
```{r}
# chatgpt: Given that we don't know which globe was tossed, we must consider both possibilities with equal weight. Hence, the total probability of observing "land" is the average of these two scenarios, represented as (Pr(Land|Earth) * Pr(Earth) + Pr(Land|Mars) * Pr(Mars)) / 2. -> 1 for mars, 0.3 for earth, divide by 2 cuz 2 planets
P_Earth <- 0.5
P_Land_given_mars <- 1
P_Land_given_earth <- 0.3
P_Land <- (P_Land_given_earth + P_Land_given_mars)/2 #because need avg of both
answer <- (P_Land_given_earth*P_Earth)/P_Land
answer
```

For the exercises below, I highly suggest you to grab a piece of paper and try to solve the problems in a 'visual' way.

2M4.
Suppose you have a deck with only three cards. Each card has two sides, and each side is either black or white. One card has two black sides. The second card has one black and one white side. The third card has two white sides. Now suppose all three cards are placed in a bag and shuffled. Someone reaches into the bag and pulls out a card and places it flat on a table. A black side is shown facing up, but you don’t know the color of the side facing down. Show that the probability that the other side is also black is 2/3. Use the counting method (Section 2 of the chapter) to approach this problem. This means counting up the ways that each card could produce the observed data (a black side facing up on the table)

There are 3 ways to get a black side (1st or 2nd card). There are 2 ways to get a black side and have the other side also be black (both from the first card). Thus the probability of pulling a card with one black side and having the other side also be black is 2/3

2M5.
Now suppose there are four cards: B/B, B/W, W/W, and another B/B. Again suppose a card is drawn from the bag and a black side appears face up. Again calculate the probability that the other side is black

now i will slightly copy the table from the solutions as that is more straightforward lol
card | ways to produce black
----------------------------
BB   | 2
BB   | 2
BW   | 1
WW   | 0

total ways to produce a black side: 5
total ways to produce a card where both sides are black: 4 (BB and BB)
thus, 4/5


2M6.
Imagine that black ink is heavy, and so cards with black sides are heavier than cards with white sides. As a result, it’s less likely that a card with black sides is pulled from the bag. So again assume there are three cards: B/B, B/W, and W/W. After experimenting a number of times, you conclude that for every way to pull the B/B card from the bag, there are 2 ways to pull the B/W card and 3 ways to pull the W/W card. Again suppose that a card is pulled and a black side appears face up. Show that the probability the other side is black is now 0.5. Use the counting method, as before.

Twice as likely to get BW than BB
Thrice as likely to get WW than BB

card | ways to produce black side | chance of getting it (prior) | posterior
-----------------------------------------------------------------------------
BB   | 2                          | 1                            | 2 (2*1)
BW   | 1                          | 2                            | 2 (1*2)
WW   | 0                          | 3                            | 0

total ways of getting a black side: 4 (2+2 from the posteriors?)
total ways of getting a card where both sides are black: 2
2/4 = 0.5